---
title: get_all()
---

It is often useful to get quick information out of an `ecopart_obj` at the start of a project. `get_all()` is a useful funciton for these quick, interactive tasks. This will return a vector of all observations for that column. Alone this data may be overwhelming, so it is a good idea to combine it with next-step summary function like `range()`, `mean()`, `max()`, etc.

If you are interested in pulling just a 

#### What is the depth range of observations?
```{r}
library(EcotaxaTools)
ecopart_example |> 
    get_all('depth_including_offset') |>
    range()
```

#### What is the average size?
When working with an `ecopart_obj`, all the size based metrics will still be in pixels, not mm. `get_all()` has an argument for converting pixels to mm. For this to work, you must provide an `ecopart_obj`, not just a `zoo_list`.
```{r}
ecopart_example |>
    get_all('esd', pixel_conv = T) |>
    mean()
```

#### Additional Info:
When using, users must specify which column they are interested in querying from the `zoo_df`. If unfamiliar with these names you can access them by getting the column names of any single `zoo_df`
```{r}
ecopart_example$zoo_files$bats361_ctd1 |> 
    names()
```